---
title: "Model 03"
output:
  word_document: default
  html_document: default
date: "2024-06-17"
---

#Importing the necessary libraries
```{r}
library(dplyr)
library(tidyr)
library(dplyr)
library(caret)
library(rpart)
library(partykit)
library(prettyR)
```

# Loading the data file from Wave 2 interviews to calculate the BMI

```{r}
load("34921-0001-Data.rda")

da34921.0001 <- da34921.0001 %>%
  mutate(
         OBESITY = case_when(
           ((WEIGHT)/(HEIGHT*HEIGHT) * 703) >= 30.000 ~ 1,
           ((WEIGHT)/(HEIGHT*HEIGHT) * 703) < 30.000 ~ 0
         )) 

obesity <- da34921.0001 %>% select(ID, OBESITY)
head(obesity)

```

# Loading and Processing the Independent Social Network Variables to calculate Bridge from WAVE 1.
```{r}

load("20541-0001-Data.rda")
load("20541-0004-Data.rda")


da20541.0001 <- da20541.0001 %>%
  select (ID, HEARN_RECODE, GENDER, AGE, RACE_RECODE, ETHGRP, COMBUILD, DEGREE_RECODE, HISPANIC, MARITLST,JOBSTAT_1, PHYSHLTH, MNTLHLTH, ATNDSERV )

```

```{r}

da20541.0001 <- da20541.0001 %>%
  mutate(DEGREE_RECODE = if_else(DEGREE_RECODE == "(-2) don't know", NA, DEGREE_RECODE),
         HEARN_RECODE = if_else(HEARN_RECODE == "(-2) don't know", NA, HEARN_RECODE),
         RACE_RECODE = if_else(RACE_RECODE == "(-2) don't know", NA, RACE_RECODE))
```

```{r}
head(da20541.0001)
nrow(da20541.0001)
```


```{r}
da20541.0004 <- da20541.0004 %>%
  group_by(ID) %>%
  filter(n() > 2) %>%
  ungroup()
```

```{r}
da20541.0004 <- da20541.0004 %>%
  pivot_longer(
    cols = starts_with("TALKFREQ"),
    names_to = "TALKFREQ",
    values_to = "FREQ"
  )

da20541.0004 <- da20541.0004 %>%
  group_by(ID) %>%
  summarize(
    BRIDGE = if_else(any(FREQ == '(0) have never spoken to each other', na.rm = TRUE), 1, 0),
    HEALTHDISCUSSIONS = if_else(any(HEALTHTALK == '(3) very likely', na.rm = TRUE), 1, 0),
    LIVEALONE = if_else(any(LIVEWITH == '(1) yes -- lives in the same household', na.rm = TRUE), 0,1))
  
head(da20541.0004)
```

```{r}
nrow(da20541.0004)
```


```{r}
modeldata <- da20541.0001 %>%
  left_join(da20541.0004, by = "ID")

modeldata <- modeldata %>%
  left_join(obesity, by = "ID")

modeldata<- na.omit(modeldata)
modeldata <- modeldata %>% select(-ID)

modeldata$BRIDGE <- as.factor(modeldata$BRIDGE)
modeldata$HEALTHDISCUSSIONS <- as.factor(modeldata$HEALTHDISCUSSIONS)
modeldata$LIVEALONE <- as.factor(modeldata$LIVEALONE)
modeldata$OBESITY <- as.factor(modeldata$OBESITY)
head(modeldata)
```

# Creating Data Partition for 70% Training Data and 30% Testing Data

```{r}
library(rpart)
library(caret)

set.seed(19032023)

index <- createDataPartition(modeldata$OBESITY, 
                             p=0.7,
                             list=FALSE, 
                             times = 1
                          )
 
modeldata.train <- modeldata[index,]
modeldata.test <- modeldata[-index,]

nrow(modeldata.train)
nrow(modeldata.test)
```

# Applying Logistic Regression on to find the association between Bridge and Obesity.

```{r}
model.lr <- glm(OBESITY ~ ., data = modeldata.train, family = "binomial")

summary.lr <- summary(model.lr)
```
# p-value for Bridge variable
```{r}
print(summary.lr)
```
```{r}
names(coef(model.lr))
```
# Odds Ratio nnd 95% Confidence Interval

```{r}
odds_ratio <- exp(coef(model.lr)["BRIDGE1"])
print(odds_ratio)
conf_int <- exp(confint(model.lr, "BRIDGE1"))
print(conf_int)
```

```{r}
predicted.prob.lr <- predict(model.lr, modeldata.test, type = "response")
predicted.obesity.lr <- ifelse(predicted.prob.lr > 0.5, 1, 0)

actual.obesity.lr <- modeldata.test$OBESITY
conf.matrix.lr <- table(Predicted = predicted.obesity.lr, Actual = actual.obesity.lr)

print(conf.matrix.lr)
```

```{r}
confusionMatrix(factor(predicted.obesity.lr), factor(modeldata.test$OBESITY), positive = as.character(1))
```


# Decision Tree 
# Conditional Inference Tree implementation using ctree

```{r}
set.seed(123)

model.dt <- ctree(OBESITY ~ ., 
                              data=modeldata.train)
plot(model.dt)

```


# Classification and Regression Tree implementation using rpart
```{r}
rpart.tree <- rpart(OBESITY ~ ., data = modeldata.train, method = "class", parms  = list(split = "information"))
rpart.tree
```

```{r}
library(rpart.plot)

rpart.plot(
  rpart.tree,
  type = 2,              # Show split labels at all nodes
  extra = 104,           # Display both percentage and number of observations
  under = TRUE,          # Show the prediction under the node
  cex = 0.6,             # Increase font size
  #tweak = 1.1,           # Fine-tune the overall size of the plot
  box.palette = "RdYlGn",# Add some color for better visualization
  compress = TRUE        # Compress the tree for better fit
)
```

```{r}
library(dplyr)

importances <- varImp(rpart.tree) %>%
  arrange(desc(Overall))

importances
```
